%
% Niniejszy plik stanowi przykład formatowania pracy magisterskiej na
% Wydziale MIM UW.  Szkielet użytych poleceń można wykorzystywać do
% woli, np. formatujac wlasna prace.
%
% Zawartosc merytoryczna stanowi oryginalnosiagniecie
% naukowosciowe Marcina Wolinskiego.  Wszelkie prawa zastrzeżone.
%
% Copyright (c) 2001 by Marcin Woliński <M.Wolinski@gust.org.pl>
% Poprawki spowodowane zmianami przepisów - Marcin Szczuka, 1.10.2004
% Poprawki spowodowane zmianami przepisow i ujednolicenie 
% - Seweryn Karłowicz, 05.05.2006
% Dodanie wielu autorów i tłumaczenia na angielski - Kuba Pochrybniak, 29.11.2016

% dodaj opcję [licencjacka] dla pracy licencjackiej
% dodaj opcję [en] dla wersji angielskiej (mogą być obie: [licencjacka,en])
\documentclass[en]{pracamgr}

% Dane magistranta:
\autor{Katarzyna Kowalska}{371053}

% Dane magistrantów:
%\autor{Autor Zerowy}{342007}
%\autori{Autor Pierwszy}{342013}
%\autorii{Drugi Autor-Z-Rzędu}{231023}
%\autoriii{Trzeci z Autorów}{777321}
%\autoriv{Autor nr Cztery}{432145}
%\autorv{Autor nr Pięć}{342011}

\title{Approximation and Parametrized Algorithms for Segment Set Cover}
\titlepl{Algorytmy parametryzowania i
trudność aproksymacji problemu pokrywania zbiorów
odcinkami na płaszczyźnie}

%\tytulang{An implementation of a difference blabalizer based on the theory of $\sigma$ -- $\rho$ phetors}

%kierunek: 
% - matematyka, informacyka, ...
% - Mathematics, Computer Science, ...
\kierunek{Computer Science}

% informatyka - nie okreslamy zakresu (opcja zakomentowana)
% matematyka - zakres moze pozostac nieokreslony,
% a jesli ma byc okreslony dla pracy mgr,
% to przyjmuje jedna z wartosci:
% {metod matematycznych w finansach}
% {metod matematycznych w ubezpieczeniach}
% {matematyki stosowanej}
% {nauczania matematyki}
% Dla pracy licencjackiej mamy natomiast
% mozliwosc wpisania takiej wartosci zakresu:
% {Jednoczesnych Studiow Ekonomiczno--Matematycznych}

% \zakres{Tu wpisac, jesli trzeba, jedna z opcji podanych wyzej}

% Praca wykonana pod kierunkiem:
% (podać tytuł/stopień imię i nazwisko opiekuna
% Instytut
% ew. Wydział ew. Uczelnia (jeżeli nie MIM UW))
\opiekun{dr Michał Pilipczuk\\
  Instytut Informatyki\\
  }

% miesiąc i~rok:
\date{June 2020}

%Podać dziedzinę wg klasyfikacji Socrates-Erasmus:
\dziedzina{ 
%11.0 Matematyka, Informatyka:\\ 
%11.1 Matematyka\\ 
%11.2 Statystyka\\ 
11.3 Informatyka\\ 
%11.4 Sztuczna inteligencja\\ 
%11.5 Nauki aktuarialne\\
%11.9 Inne nauki matematyczne i informatyczne
}

%Klasyfikacja tematyczna wedlug AMS (matematyka) lub ACM (informatyka)
\klasyfikacja{D. Software\\
  D.127. Blabalgorithms\\
  D.127.6. Numerical blabalysis}

% Słowa kluczowe:
\keywords{set cover, geometric set cover, FPT, W[1]-completeness,
APX-completeness, PCP theorem, NP-completeness}

% Tu jest dobre miejsce na Twoje własne makra i~środowiska:

\newcommand{\points}{\mathcal{C}}
\newcommand{\sets}{\mathcal{P}}
\newcommand{\sol}{\mathcal{R}}
\newcommand{\then}{\Rightarrow}

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage[nospace, noadjust]{cite}
\usepackage{lineno}
\usepackage{enumitem}
\usepackage{amsthm}
\usepackage{mathtools}  
\usepackage{makecell}
\usepackage{tikz}
\usetikzlibrary{calc,math}
\linenumbers

\mathtoolsset{showonlyrefs}  

\theoremstyle{plain}
\newtheorem{claim}{Claim}[chapter]
%\newtheorem{defi}{Definition}[section]
\newtheorem{tw}{Theorem}[chapter]
\newtheorem{lemma}{Lemma}[chapter]
\newtheorem{corollary}{Corollary}[chapter]
\newtheorem{remark}{Remark}[chapter]

\theoremstyle{definition}
\newtheorem{defi}{Definition}[chapter]

\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}


% koniec definicji

\begin{document}
\maketitle

%tu idzie streszczenie na strone poczatkowa
\begin{abstract}
  The work presents a study
  of different geometric set cover problems.
  It mostly focuses on segment set cover
  and its connection to the polygon set cover.
\end{abstract}

\tableofcontents
%\listoffigures
%\listoftables

\chapter{Introduction}

The Set Cover problem is one of the most common NP-complete problems.
[tutaj referencja]
We are given a family of sets and have to choose the smallest
subfamily of these sets that cover all their elements.
This problem naturally extends to settings
were we put different weights on the sets
and look for the subfamily of the minimal weight.
This problem is NP-complete even 
without weights and if we put
restrictions on what the sets can be.
One of such variants is Vertex Cover problem,
where sets have size 2 (they are edges in a graph).

In this work we focus on another such variant where the sets correspond
to some geometric shapes and
only some points of the plane have to be covered.
When these shapes are rectangles with edges parallel
to the axis, the problem can be proven to
be W[1]-complete (solution of size $k$ cannot be found
in $n^o(k)$ time),
APX-complete (for suffciently small $\epsilon > 0$, the problem
does not admit $1+\epsilon$-approximation scheme)
[refrencje].

Some of these settings are very easy.
Set cover with lines parallel to one of the axis
can be solved in polynomial time.

There is a notion of $\delta$-expansions,
which loosen the restrictions on geometric set cover.
We allow the objects to cover the points
after $\delta$-expansion and compare
the result to the original setting.
This way we can produce both FPT and EPTAS
for the rectangle set cover with $\delta$-extensions
[referencje].



\paragraph{Our contribution.}
In this work, we prove that unweighted geometric set cover
with segments is fixed parameter tractable (FPT).

Moreover, we show that geometric set cover with segments
is APX-complete for unweighted axis-parallel segments,
even with 1/2-extensions.
So the problem for very thin rectangles
also cannot admit PTAS.
Therefore, in the efficient polynomial-time approximation scheme (EPTAS)
for \textit{fat polygons} by \cite{harpeled12},
the assumption about polygons being fat is necessary. 

Finally, we show that geometric set cover with weighted segments in
3 directions is W[1]-complete.
However, geometric set cover with weighted segments is FPT if we allow
$\delta$-extension.

This result is especially interesting,
since it's counter-intuitive that
the unweighed setting is FPT and the weighted
setting is W[1]-complete.
Most of such problems (like vertex cover or [wiecej przykladow])
are equally hard in both weighted and unweighted settings.

\input{definitions.tex}
\chapter{APX-hardness geometric set cover problem}
\input{apxcomplete.tex}
\chapter{Fixed-parameter tractable algorithm for geometric set cover problem}
\input{segments.tex}
\input{fpt_weighted_segments.tex}
\input{w1_completeness.tex}

\chapter{Geometric Set Cover with lines}
\section{Lines parallel to one of the axis}
When $\mathcal{R}$ consists only of lines parallel to
one of the axis, the problem can be solved in
polynomial time.

We create bipartial graph $G$ with node for every line on the input
split into sets: $H$ -- horizontal lines and $V$ -- vertical lines.
If any two lines cover the same point from $\mathcal{C}$, then
we add edge between them.

Of course there will be no edges between nodes inside $H$,
because all of them are pararell and if they share 
one point, they are the same lines. Similar argument for $V$.
So the graph is bipartial.

Now Geometric Set Cover can be solved with Vertex Cover on graph $G$.
Since Vertex Cover (even in weighted setting) 
on bipartial graphs can be solved in polynomial time.

Short note for myself just to remember how to this in polynomial time:

Non-weighted setting - Konig theorem + max matching

Weighted setting - Min cut in graph of $\neg A$ or $\neg B$
(edges directed from $V$ to $H$)

\section{FPT for arbitrary lines}
You can find this is Platypus book.
We will show FPT kernel of size at most $k^2$.

(Maybe we need to reduce lines with one point/points with one line).

For every line if there is more than $k$ points on it,
you have to take it. At the end, if there is more than $k^2$ points,
return NO.
Otherwise there is no more than $k^4$ lines.

In weighted settings among the same lines with different weights
you leave the cheapest one and use the same algorithm.

\section{APX-completeness for arbitrary lines}
We will show a reduction from Vertex Cover problem.
Let's take an instance of the Vertex Cover problem for graph $G$.
We will create a set of $|V(G)|$ pairwise non-pararell lines,
such that no three of them share a common point.

Then for every edge in $(v, w) \in E(G)$
we put a point on crossing of lines for vertices $v$ and $w$.
They are not pararell, so there exists exactly one such point
and any other line do not cover this point (any three of them do not
cross in the same point).

Solution of Geometric Set Cover for this instance would yield
a sound solution of Vertex Cover for graph $G$.
For every point (edge) we need to choose at least one of
lines (vertices) $v$ or $w$ to cover this point.

Vertex Cover for arbitrary graph is APX-complete,
so this problem in also APX-complete.

\section{2-approximation for arbitrary lines}
Vertex Cover has an easy 2-approximation algorithm,
but here very many lines can cross through
the same point, so we can do $d$-approximation,
where $d$ is the biggest number of lines crossing through the same point.
So for set where any 3 lines do not cross in the same point
it yields 2-approximation.

The problematic cases are where through all points
cross at least $k$ points and all lines have at least $k$ points on them.
It can be created by casting $k$-grid in $k$-D space on 2D space.

Greedy algorithm yields $\log |\mathcal{R}|$-approximation,
but I have example for this for bipartial graph and
reduction with taking all lines crossing through some point
(if there are no more than $k$) would solve this case.
So maybe it works.

Unfortunaly I have not done this :(

I can link some papers telling it's hard to do.

\section{Connection with general set cover}
Problem with finite set of lines with more dimensions
is equivalent
to problem in 2D, because we can project
lines on the plane which is not perpendicular
to any plane created by pairs of
(point from $\mathcal{C}$, line from $\mathcal{P}$).

Of course every two lines have at most one common point,
so is every family of sets that have at most one point
in common equivalent to some geometric set cover with lines?

No, because of Desargues's theorem.
Have to write down exactly what configuration is banned.


\chapter{Geometric Set Cover with polygons}
\section{State of the art}

Covering points with weighted discs admits PTAS \cite{li}
and with fat polygons with $\delta$-extensions with unit weights
admits EPTAS \cite{harpeled12}.

Although with thin objects, even if we allow $\delta$-expansion,
the Set Cover with rectangles
is APX-complete (for~$\delta = 1/2$),
it follows from APX-completeness for segments with $\delta$-expansion
in Section \ref{section:segment_apx}.

Covering points with squares is W[1]-hard \cite{marx05}.
It can be proven that assuming $SETH$,
there is no $f(k)\cdot(|\points|+|\sets|)^{k-\epsilon}$ time algorithm
for any computable function $f$ and $\epsilon >0$ that decides if there
are $k$ polygons in $\sets$ that together cover $\points$,
\textit{Theorem 1.9} in \cite{voronoi}.








\chapter{Conclusions}
We do not know FPT for axis-parallel segments without $\delta$-extensions.

\bibliographystyle{apalike}
\bibliography{bibl}

\end{document}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% coding: latin-2
%%% End:
