\chapter{Preliminaries}

In this chapter we present some basic definitions that
will be used later.

\section{$\GeometricSetCover$}
\label{section:def:geometric__set_cover}
Whenever speaking about $\GeometricSetCover$,
we consider it in the 2-dimensional plane.

In the $\GeometricSetCover$ problem we are are given
$\sets$ --- a set of objects, which are connected
subsets of the plane and $\points$ --- a set of points in the plane.
The task is to choose $\sol \subseteq \sets$ such that
every point in $\points$ is inside some object from $\sol$
and $|\sol|$ is minimized. We will mostly consider the case where
$\sets$ consists of segments in the plane.

In the weighted setting, there is some given weight function
$f : \sets \rightarrow \mathbb{R^+}$
and we would like to find a solution $\sol$
that minimizes $\sum_{R \in \sol} f(R)$.

\begin{defi}
A segment is \textbf{axis-parallel} if it lies on a line that is
either horizontal $y = c$ or vertical $x = c$.
\end{defi}

\begin{defi}
	A line is \textbf{right-diagonal} if it is
	described by the linear function $x + y = d$ for some $d \in \mathbb{R}$.
	A segment is \textbf{right-diagonal} if its
	direction is a right-diagonal line.
\end{defi}

\section{Parameterization}

In the parameterized setting of the $\GeometricSetCover$
for a given $k$,
our task is to either find a solution $\sol$ such that $|\sol| \le k$
or decide that there is no such solution.

\begin{defi}
A \textbf{fixed-parameter (FPT)} algorithm 
for a problem with parameter $k$ and instance size $n$
is an algorithm running in time $f(k) \cdot n^c$
for some constant $c$ and some computable function $f$.
\end{defi}

\begin{defi}
\label{definition:cnf}
Boolean formula is in \textbf{conjunctive normal form (CNF)} if
it is a conjunction of one or more formulas,
which are disjunction of literals.
\textbf{$k$-CNF} formula is a CNF formula, where
every disjunction consists of at most $k$ literals.
\end{defi}

\begin{defi}
\textbf{\textsc{$k$-SAT}} problem is 
a Boolean satisfiability problem of $k$-CNF formulas.
Given $k$-CNF formula, one must answer if there
exists any variable assignment that satisfies the formula.
\end{defi}

\begin{defi}
For $k \ge 3$, let us define $S_k$ as the set of constants $\sigma$
such that there exists an algorithm solving \textsc{$k$-SAT} running in time
$2^{\sigma n} \cdot n^{\mathcal{O}(1)}$.
Let $s_k$ be the infimum  of the set $S_k$.

\textbf{Exponential Time Hypothesis (ETH)} asserts
that $s_3 > 0$. This conjecture implies that
there does not exist an algorithm solving \textsc{3-SAT}
running in time $2^{o(n)}$.
\end{defi}

The definition of a W[1]-hard problem and W hierarchy,
can be found in Chapter 13.3 of \cite{platypus_book}.
When proving that a problem is W[1]-hard,
we are going to use Theorem~\ref{grid_tiling_w1_hard}
(W[1]-hardness of \textsc{Grid Tiling}),
which was proved in \cite{marx_grid_tiling}.

\section{Approximation}

Let us recall some definitions related to optimization problems.

\newcommand{\opt}{\mathsf{opt}}

\begin{defi}
A \textbf{polynomial-time approximation scheme (PTAS)}
for a minimization problem $\Pi$
is a family of algorithms $\cal{A}_\epsilon$ for
every $\epsilon > 0$
such that $\cal{A}_\epsilon$ takes an instance $I$ of~$\Pi$
and in polynomial time
finds a solution that is within a factor
of ($1+\epsilon$) of being optimal.
This means that the reported solution has weight at most
$(1+\epsilon)\opt(I)$, where $\opt(I)$ is the weight
of an optimal solution to $I$.
\end{defi}

\begin{defi}
A problem $\Pi$ is \textbf{APX-hard} if assuming P $\neq$ NP,
there exists $\epsilon > 0$
such that there is no polynomial-time $(1+\epsilon)$-approximation algorithm
for $\Pi$.
\end{defi}

\section{$\delta$-extension}
\label{section:def:delta_extension}

Another idea presented here, which can be utilized only when considering
the problems with geometric objects,
is $\delta$-extension.
We define it specifically for the $\GeometricSetCover$ problem
with convex centre-symmetric objects.

Intuitively, we consider a problem with slightly larger objects,
which makes the instance more permissive.
However, we aim to find a solution that
is not larger than the
optimum solution to the original problem,
so this is substantially easier than just
solving the problem for the larger objects.
It may even be the case
that we are able to find a solution
of size smaller than the optimum solution
to the original problem.

Formal definition of $\delta$-extended objects
is present in Definition
\ref{definition:delta_extension}.

The $\GeometricSetCover$ with $\delta$-extension
is a version of $\GeometricSetCover$ with
the following modifications.
\begin{itemize}
\item We need to cover all the points in $\points$
by selecting objects from $\{P^{+\delta} : P \in \sets\}$ (which always 
include no fewer points than the objects
before $\delta$-extension).
\item We look for a solution that is not larger than the optimum
solution to the original problem.
Note that it does not need to be an optimal solution in
the modified problem.
\end{itemize}
Formally, we have the following.

\begin{defi}
The \textbf{$\GeometricSetCover$ problem
with $\delta$-extension} is the problem where for an input instance
$I=(\sets, \points)$ of $\GeometricSetCover$,
the task is to output a solution $\mathcal{R} \subseteq \sets$
such that the~$\delta$-extended set
$\{ R^{+\delta} :  R \in \mathcal{R} \}$ covers $\points$
and is not larger than the optimal solution to the~problem without
extension, i.e.~$|\mathcal{R}| \le |opt(I)|$.
\end{defi}

At last, we formulate a definition of the
polynomial-time approximation scheme (PTAS)
for a problem with $\delta$-extension.

\begin{defi}
A \textbf{PTAS for $\GeometricSetCover$
with $\delta$-extension} is a family of algorithms
$\{\mathcal{A}_{\delta, \epsilon}\}_{\delta, \epsilon > 0}$ that
each takes as an input instance $I=(\sets, \points)$
of \textsc{Geometric Set} \textsc{Cover} where objects are centre-symmetric and convex,
and in polynomial-time outputs a solution $\mathcal{R} \subseteq \sets$
such that the $\delta$-extended set
$\{ R^{+\delta} :  R \in \mathcal{R} \}$ covers $\points$
and is within a $(1+\epsilon)$ factor of the optimal
solution to this problem without
extension, i.e.~$(1+\epsilon)|\mathcal{R}| \le |opt(I)|$.
\end{defi}

\section{$\WeightedGeometricSetCover$}

In this thesis we also consider a $\WeightedGeometricSetCover$ problem,
which is a combination
of the weighted and parameterized setting described in 
Section \ref{section:def:geometric__set_cover}.
We already argued in the introduction
that there is no consensus of how it is defined, but when we discuss the
weighted parameterized setting we will consider the following
definition. There is a given weight function
$f : \sets \rightarrow \mathbb{R^+}$
and we would like to find a solution $\sol$
such that $|\sol| \le k$
and $\sum_{R \in \sol} f(R)$ is minimum possible among such sets $\sol$.

\begin{defi}
The \textbf{$\WeightedGeometricSetCover$ problem
with $\delta$-extension} is the problem where for an input instance
$I=(\sets, \points, f)$ of \textsc{Weighted Geometric Set} \textsc{Cover},
the task is to output a solution $\mathcal{R} \subseteq \sets$
such that the~$\delta$-extended set
$\{ R^{+\delta} :  R \in \mathcal{R} \}$ covers $\points$
and it has weight not larger than the optimal solution to the~problem without
extension, i.e.~$\sum_{R \in \mathcal{R}} f(R) \le |\opt(I)|$.
\end{defi}

We also consider weighted parameterized setting with $\delta$-extension,
which we formally define below.

\begin{defi}
The \textbf{$\WeightedGeometricSetCover$ problem
with $\delta$-extension parameterized by the size of a solution}
is a problem where for an input instance
${I=(\sets, \points, f, k)}$ of $\WeightedGeometricSetCover$
parameterized by the size of a solution $k$,
the task is to output a solution $\mathcal{R} \subseteq \sets$
such that the~$\delta$-extended set
$\{ R^{+\delta} :  R \in \mathcal{R} \}$ covers $\points$,
uses no more than $k$ sets, i.e. $|\sol| \le k$
and it has weight not larger than the optimal solution to the~problem without
extension, i.e.~$\sum_{R \in \mathcal{R}} f(R) \le |\opt(I)|$.
\end{defi}
