In this chapter we show fixed-parameter tractable algorithms
for the geometric set cover problem in~two different settings.
Section \ref{section:fpt_unweighted} shows 
a fixed-parameter tractable algorithm for geometric set cover with unweighted segments.
The remainder of the chapter presents
a fixed-parameter tractable algorithm for geometric set cover with weighted segments
with $\delta$-extension.
We show an algorithm for the setting with $\delta$-extension,
because the original problem with weights is W[1]-hard,
as we show in Chapter $\ref{chapter:w1_hard}$.

We start with a shared definition for this problem.
We define \textit{extreme points} for a set of~collinear points.

\begin{defi}
	For a set of collinear points $C$ in the plane,
	\textbf{extreme points} of $C$ are the endpoints
	of the smallest segment that covers all points from set $C$.
	
	If $C$ consists of one point or is empty, then
	there are 1 or 0 extreme points respectively.
\end{defi}

\section{Fixed-parameter tractable algorithm for $\SegmentSetCover$}
\label{section:fpt_unweighted}
In this section we consider fixed-parameter tractable
algorithms for unweighted geometric set cover with segments.
The setting where segments are required to be axis-parallel
(or limited to a constant number of directions) has a trivial FPT algorithm.
We present an FPT algorithm for geometric set cover
with unweighted segments, where segments are in arbitrary directions.

\subsection{Axis-parallel segments}
\begin{tw}
	\textbf{(FPT for segment cover with axis-parallel segments)}.
	There exists an algorithm that given a family $\sets$ of
	axis-parallel segments,
	a set of points $\points$
	and a parameter $k$,
	runs in time $\mathcal{O}(2^k)$,
	and outputs a solution $\sol \subseteq \sets$
	such that $|\sol| \le k$ and $\sol$ covers all points in~$\points$,
	or determines that such a set $\sol$ does not exist.
\end{tw}

\begin{proof}
We show an $\mathcal{O}(2^k)$-time branching algorithm.
In each step, the algorithm selects a point $a$ which is not yet covered,
branches to choose one of the two directions, and greedily chooses
a segment $a$ in that direction to cover.
This proceeds until either all points are covered or $k$~segments are chosen.

Let us take
the point $a=(x_a,y_a)$ which is the smallest 
among points that are not yet covered
in the lexicographic ordering
of points in $\mathbb{R}^2$.
We need to cover $a$ with some of~the~remaining segments.

Branch over the choice of one of the coordinates ($x$ or $y$);
without loss of generality, let us assume we chose $x$.
Among the segments lying on line $x = x_a$,
we greedily add to~the~solution the~one that covers the most points.
As $a$ was the smallest in the lexicographical order,
all points on the line $x = x_a$ have the $y$-coordinate larger than $y_a$.
Therefore, if we denote the~greedily chosen segment as $s$,
then any other segment on the line $x = x_a$ that covers $a$ can only
cover a subset of points covered by $s$.
Thus, greedily choosing $s$ is optimal.

In each step of the algorithm we add one segment to the solution,
thus the recursion can be stopped at depth $k$.
If no branch finds a solution, then this means
that a solution of size at most $k$ does not exist.
\end{proof}

Note that the same algorithm can be used for segments in $d$ directions,
where we branch over $d$ choices of directions, and it runs in complexity $\mathcal{O}(d^k)$.

\subsection{Segments in arbitrary directions}
\label{segments_in_arbitrary_direction}
In this section we consider the setting where segments are not constrained
to a constant number of directions. 
We present a fixed-parameter tractable algorithm,
parameterized by the size of the solution.

\segmentCoverFpt*

We will need the following lemmas proving properties of any
instance of the problem.

\begin{lemma}
   \label{fpt_reduction}
   Given an instance $(\sets, \points)$ of the segment cover problem,
   without loss of generality we can assume that
   no segment covers a superset of what another segment covers.
   That~is, for any distinct $A, B \in \sets$, we have
   $A \cap \points \not \subseteq B \cap \points$ and $A \cap \points \not \supseteq B \cap \points$.
\end{lemma}   
   
\begin{proof}
Assume towards a contradiction that there is an instance  $(\sets, \points)$,
and two distinct subsets of $\sets$,
$A, B$, such that $A \cap \points \subseteq B \cap \points$.

We construct a set $\sets' := \sets - \{A\}$.
We prove that for any solution $\sol$ of $(\sets, \points)$,
we can construct a~solution $\sol' \subseteq \sets'$,
such that $|\sol'| \le |\sol|$.
Let us take any solution $\sol$ of $(\sets, \points)$.
If $A \in \sol$, then $\sol' := \sol \cup \{B\} - \{A\}$,
otherwise $\sol' := \sol$.
Let us consider the case when $A \in \sol$,
because the other case is trivial.
Since $A \cap \points \subseteq B \cap \points$,
then $\sol \cup \{B\} - \{A\}$
covers any point from $\points$ that was covered by $\sol$.
Also, $|\sol \cup \{B\} - \{A\}| \le |\sol|$.
\end{proof}

\begin{lemma}
	\label{fpt_long_lines}
	Given an instance $(\sets, \points)$
	of the segment cover problem 
	transformed by Lemma~\ref{fpt_reduction},
	if there exists a line $L$ with at least
	$k+1$ points on it, then there exists a subset $A \subseteq \sets$,
	of size at most $k$,
	such that every solution $\sol$ with $|\sol| \le k$
	satisfies $|A \cap \sol| \ge 1$.
	Moreover, such a subset can be found in~polynomial time.
\end{lemma}
\begin{proof}
Let us enumerate the points from $\points$ that lie on $L$ as $x_1, x_2, \ldots, x_t$
in the order in which they appear on $L$.
Our proposed set is defined as:
$$A := \left\{ \text{segment collinear with } L \text{ that covers } x_i
\text{ and does not cover } x_{i-1} : i \in \{1, \ldots, k\}\right\},$$
where for $i = 1$ we just take a segment that covers $x_1$.
If such a segment does not exist for any point $x$
as above, then $x$ does not give rise to any segment in $A$.

We prove the lemma by contradiction. Let us assume that there
exists a~solution $\sol$ of size at most $k$ such that $\sol \cap A = \emptyset$.


Let $\sol_L$ be the set of segments from $\sol$ that are collinear with $L$.

Every segment that is not collinear with $L$ can cover at~most one of
the points that lie on~this line.
Hence, if $\sol_L$ was empty, then
$\sol$ would cover at most $k$ points on line $L$,
but $L$ had at least $k+1$ different points from $\points$ on it.

Therefore, we know that $\sol_L$ is not empty
and $|\sol - \sol_L| \le k-1$.
Segments from $\sol - \sol_L$ can cover at most $k-1$
points among $\{x_1, x_2, \ldots, x_k\}$, therefore at least
one of these points must be covered by segments from $\sol_L$.
We take the leftmost point from $\{x_1, x_2, \ldots, x_k\}$ that is
covered in $\sol_L$ and name it $a$.
After the transformation from Lemma \ref{fpt_reduction},
in $\sol$ there is only one segment
that starts in $a$ and is collinear with $L$,
therefore this segment must be in both $\sol$ and $A$.
This contradiction concludes the proof that $|A \cap \sol| \ge 1$
for any solution $\sol$ of size at most $k$.
\end{proof}

We are now ready to prove Theorem \ref{segment_cover_fpt}.

\begin{proof}[Proof of Theorem \ref{segment_cover_fpt}.]
We will prove this theorem by presenting a branching algorithm that
works in desired complexity. It first branches over the
choice of segments to cover the lines with \textit{many} points
and then solves a small instance (where every line has at most $k$ points)
by checking all possible solutions.

\subparagraph{Algorithm.}
We present a recursive algorithm. Given an instance of the problem:

\begin{enumerate}[label={(\arabic*)}]
\item Use Lemma \ref{fpt_reduction} to remove some redundant segments from our instance.
\item If there exists a line with at least $k+1$ points from $\points$,
we branch over the choice of adding to~the~solution
one of~the~at~most $k$ possible segments
provided by Lemma \ref{fpt_long_lines}; name this segment $s$
and name the set of points from $\points$ that lie on $s$ as $S$.
By recursion, we find a~solution $\sol$
for the instance $(\points - S, \sets - \{s\})$,
and parameter $k-1$. We return $\sol \cup \{s\}$.
Note that if Lemma \ref{fpt_long_lines} returned $\emptyset$,
then we respond \texttt{NO}.
\item If every line has at most $k$ points on it and $|\points| > k^2$,
then answer \texttt{NO}.
\item If $|\points| \le k^2$, solve the problem by brute force:
check all subsets of $\sets$ of size at most $k$.
\end{enumerate}

\subparagraph{Correctness.}

Lemma \ref{fpt_long_lines} proves that at least one segment that we
branch over in (1) must be present in every solution $\sol$ with $|\sol| \le k$.
Therefore, the recursive call can find a~solution, provided there exists one.

In (2) the answer is no, because every line covers no more than $k$ points
from $\points$, which implies the same about every segment from $\sets$.
Under this assumption
we can cover only $k^2$ points with a solution of size $k$, which is less
than $|\points|$.

Checking all possible solutions in (3) is trivially correct.


\subparagraph{Complexity.}

In the leaves of the recursion we have $|\points| \le k^2$, so $|\sets| \le k^4$,
because every segment can be uniquely identified by the two extreme points it covers
(by Lemma \ref{fpt_reduction}). Therefore, there are $\binom{k^4}{k}$
possible solutions to check, each can be checked in time $\mathcal{O}(k|\points|)$.
Thus, (3) takes time $k^{\mathcal{O}(k)}$.


In this branching algorithm our parameter $k$ is decreased with every
recursive call, so we have at most $k$ levels of recursion with
branching over $k$ possibilities. Candidates to branch over
can be found on each level in time $\mathcal{O}((|\points|\cdot|\sets|)^{\mathcal{O}(1)})$.

Reduction from Lemma \ref{fpt_reduction} can be implemented
in time $\mathcal{O}((|\points|\cdot|\sets|)^{\mathcal{O}(1)})$.

It follows that the overall complexity
is $\mathcal{O}(((|\points|\cdot|\sets|)^{\mathcal{O}(1)}) \cdot k^{\mathcal{O}(k)})$
\end{proof}

