\section{Weighted segments}
\subsection{FPT for weighted segments with $\delta$-extensions}
\begin{tw}[FPT for weighted segment cover with $\delta$-extensions]{
	There exists an algorithm $\mathcal{A}$ that given a family $\sets$ of
	$n$ weighted segments (in any direction),
	a set of $m$ points $\points$
	and a parameter $k$,
	runs in time $f(k) \cdot (nm)^c$ for some computable function $f$ and a constant $c$,
	and outputs a set $\sol \subseteq \sets$
	such that $|\sol| \le k$ and $\sol^{+\delta}$ covers all points in $\points$.
}\end{tw}


To solve this problem we will introduce a lemma about choosing
\textit{good} subset of points.

\begin{defi}
	For a segment $s$ and a set
	of points $C$ that lay on $s$,
	a set $A \subseteq C$ is \textbf{$(k,\delta)$-good} 
	if any set of segments $R$ that covers set $A$
	such that $|R| \le k$, it holds $R^{+\delta}$ covers $C$.
\end{defi}

\begin{lemma}
	\label{good_set_exists}
	For any segment $s$, set of points $C$, $\delta > 0$ and $k \ge 1$,
	there exists a $(k,\delta)$-good set of size at most $f(k, \delta)$
	for some computable function $f$.
\end{lemma}

\begin{proof}
We prove this for a fixed $\delta < 1$
by induction over $k$ for any segment $s$ and set of points $C$.

\subparagraph{Inductive hypothesis}
For any segment $s$ and a set of points $C$ that lay on $s$,
there exists a set $A$,
that is $(l, \delta)-good$ for every $1 \le l \le k$ and $|A| < f(\delta, k)$
for some computable function $f$.

Set $A$ also includes ends of segment $s$ for convinience.

\subparagraph{Base case for $k = 1$}
It is sufficient that $A$ consists of 2 points, end of segment $s$.

If they are covered with one segment, it must be a segment 
that includes $s$, so it covers set $C$.

\subparagraph{Inductive step}
Assuming inductive hypothesis for any $s$ and any set of points $C$
that lay on $s$ and for $k$, we will prove hypothesis for $k+1$.


We define $\lceil3/\delta\rceil$ subsegments of $s$ in the following way.
We split $s$ into $\lceil3/\delta\rceil$ segments $v_i$
such that $|v_i| = \frac{|s|\delta}{3}$. (pewnie sufit tutaj)

$C_i$ is a subset of $C$ such that they lay on $v_i$.

$t_i$ is a segment connecting leftmost and rightmost point in $C_i$
(it might by degenerated segment if $|C_i| = 1$ or it might be empty
if $C_i$ is empty).

TODO: Add a picture with $v_i$ and $t_i$ here

We use inductive hypothesis to choose $(k, \delta)$-good sets $A_i$
for segments $t_i$ and sets $C_i$. If $|C_i| \le 1$, then $A_i = C_i$
and it's still a $(k, \delta)$-good set.

Then we define $A = \bigcup_i=1^\lceil3/\delta\rceil A_i$.
It includes ends of $s$, because they are in sets $A_1$ and $A_\lceil3/\delta\rceil$.

\subparagraph{Proof that $A$ is $(k, \delta)$-good for $s$ and $C$}
Let us take any cover of $A$ with $k+1$ segments and name it $\sol$.

For every segment $t_i$, if there exists a segment $x$ from $\sol$ 
such that it is disjoint with $t_i$,
then we have a cover of $A_i$ with at most $k$
segments using $\sol - \{x\}$.
Since $A_i$ is $(k, \delta)$-good for $t_i$ and $C_i$,
then $(\sol - \{x\})^{+\delta}$ covers $C_i$.

If there doesn't exist such segment, then
there exists a segments $t_i$ such that all $k+1$ segments that cover
$A_i$ intersect with $t_i$. (Note: There exists only one such segment $t_i$).
By definition ends of $s$ are in $A_i$, so there must exist
segments starting in the ends of $s$ and ending somewhere in $t_i$.
Let us name these two segments $y$ and $z$. It follows that:
$|y| + |z| + |t_i| \le |s|$.
Since $|t_i| \le |v_i| = \frac{|s|\delta}{3}$,
therefore $max(|y|, |z|) > |s|(1-\frac{\delta}{3})/2$.


TODO: Add a picture with such segments here

After $\delta$-extension, the longer of these segments will
lengthen both ways by at least:
$$|s|(1-\frac{\delta}{3})/2\delta = |s|\delta(1/2 - \delta/6) > |s|\delta(1/2 - 1/6) = \frac{|s|\delta}{3}$$.

Therefore the longer of segments $y$ and $z$ will cover all segments $t_i$
after $\delta$-extension, therefore $\sol^{+\delta}$ covers $C_i$.

Since $C = \bigcup_{i=1}^\lceil3/\delta\rceil C_i$,
then $\sol^{+\delta}$ covers $C$.

\end{proof}

\begin{lemma}
	\label{kernel_wscps}
	\textbf{Kernel in WSCPS}.
	TODO: formulate it properly
	
	For segment cover, there is a kernel of size $f(k)$
	in WSCPS.
\end{lemma}

\begin{claim}
If there are more than $k$ lines with
at least $k+1$ points on them, then 
they can't be covered with $k$ segments.
\end{claim}

\begin{claim}
If there is more than $k^2$ points
that don't lie on any line with more than $k$
points on it, then they can't be covered with $k$ segments.
\end{claim}

\begin{claim}
For every long line $L$ (with more than $k$ points on them)
we can choose $f(k)$ points on them,
that if we cover all of these points with at most $k$
segments, then the rest of the points
with $\delta$-extensions will be covered by
segments in the direction of line $L$.
\end{claim}


\paragraph{Proof of Lemma \ref{kernel_wscps}.}
After applying the previous lemmas,
we have at most $k^2 + k\cdot f(k)$
points that can be covered in any direction
and for the rest of the points we can draw 
at most $k\cdot f(k)$ segments along
their respective long lines that have to be covered
by segments after $\delta$-extensions.

Then we extend every available segment by $\delta$-extension
and we achieve the kernel in WSCPS for this instance of problem.


\begin{lemma}
	\label{covered_spaces_lemma}
	If all the points are covered with $k$ segments and
	the biggest $2(1 + 1/\delta)^{k+1}$ spaces between points
	are filled, the whole segment is filled after
	$\delta$-extensions of these segments.
\end{lemma}
\paragraph{Proof.}
	Let's name the $2(1 + 1/\delta)^{k+1}$-st biggest space
	between points as $y$.
	We have guarantee that all segements of length $x > y$
	are covered without $\delta$-extensions.
	
	Let's take one space between points that is not covered
	before $\delta$-extension and we will prove
	it will be covered after $\delta$-extensions.
	Let's assume it isn't.
	
	This space has length $x$. Since it's uncovered, $x \le y$.
	
	Let's take side where the sum of lengths of segments
	covering the points is greater (left or right).
	Without loss of generality, let us assume it's right.
	
	There are at most $k$ segments to the right of this space
	between points. Name their lengths $l_1, l_2 \ldots l_k$.
	If the point is covered in the other direction,
	the segment is degenerated to the point and $l_i = 0$.
	Name the space between endpoints of $l_i$ and $l_{i+1}$ -- $x_i$.
	Of course, $x_i$ is uncovered space between two points,
	therefore $x_i \le y$.
	
	TUTAJ BEDZIE PEWNIE RYSUNEK Z TYMI SUPER RZECZAMI DO PRZERW
	
	Let's write equations meaning that $i$-th segment
	doesn't cover space $x$ after $\delta$-expansion.
	
	$$l_1\delta < x \le y \then l_1 < y/\delta$$
	$$l_2\delta < x + l_1 + x_1 < 2y + y/\delta \then l_2 < 2y/\delta + y/\delta^2$$
	$$l_3\delta < x + l_1 + x_1 + l_2 + x_2 < 3y + 3y/\delta + y/\delta^2 \then l_3 < 3y/\delta +  3y/\delta^2 + y/\delta^3$$
	
	From this we can "guess" induction $l_i < y((1+1/\delta)^i - 1)$
	
	Trivailly for $l_1 < y/\delta$.
	
	Assume that for all $j < i$: $$l_j < y((1+1/\delta)^j - 1)$$.
	
	$l_i\delta < x + \sum_{j = 1}^{i-1}(l_j + x_j)
	< iy \sum_{j = 1}^{i-1}l_j
	< iy + \sum{j=1}^{i-1}y((1+1/\delta)^j - 1)
	= iy - (i-1)y + \sum{j=1}^{i-1}y(1+1/\delta)^j
	= y(1 + \sum_{j = 1}^{i-1}(1+1/\delta)^j)
	= y(2 + \sum_{j = 1}^{i-1}(1+1/\delta)^j - 1)
	= y(\sum_{j = 0}^{i-1}(1+1/\delta)^j - 1)
	= y((1+1/\delta)^i / (1 - (1+1/\delta)) - 1)
	= y((1+1/\delta)^i\delta - 1)
	< y((1+1/\delta)^i\delta - \delta)$
	
	Of course we also know that (since we have chosen the side with greater sum of the width of segments):
	$$\sum_{i=1}^{k} l_i \ge 1/2 \cdot y \cdot 2(1 + 1/\delta)^{k+1} =  y \cdot (1 + 1/\delta)^{k+1}$$
	
	But 
	$\sum_{i=1}^{k} l_i
	< \sum_{i=1}^{k} y((1+1/\delta)^i - 1)
	= y((1+1/\delta)^{k+1} / (1-(1+1/\delta)) - k)
	= y((1+1/\delta)^{k+1}\delta - k)
	< y(1+1/\delta)^{k+1}$
	
	Therefore the space must have been covered after $\delta$-expansions.

\subsection{W[1]-completeness for weighted segments in 3 directions}

\begin{tw}
	\textbf{W[1]-completeness for weighted segments in 3 directions}.
	Consider the problem of covering a set $\points$ of points
	by selecting $k$ axis-pararell or right-diagonal weighted segments
	with weights
	from a set $\sets$ with minimal weight.
	Assuming ETH, there is no algorithm for this
	problem with running time
	$f(k)\cdot(|\points| + |\sets|)^{o(\sqrt(k))}$
	for any computable function $f$.
\end{tw}

We will show reduction from grid tiling problem.


Let's have an instance of grid tiling problem -- size of the
gird $k$, number of elements available $n$
and $k^2$ sets of available pairs in every tile
$S_{i, j} \subseteq \{1,n\} \times \{1,n\}$.

\paragraph{Construction.}
We construct a set $\sets$ of segments and a set $\points$
of points.

First let's choose any ordering of $n^2$ elements
$\{1,n\} \times \{1,n\}$ and name this sequence $a_1 \ldots a_{n^2}$.

$$match_v(i, j) \iff
a_i = \{x_i, y_i\} \land a_j = \{x_j, y_j\} \land x_i = x_j$$

$$match_h(i, j) \iff
a_i = \{x_i, y_i\} \land a_j = \{x_j, y_j\} \land y_i = y_j$$


\subparagraph{Points.}

Define points:
	$$h_{i, j, t} = (j \cdot (n^2+1) + t, (n^2+1) \cdot i)$$
	$$v_{i, j, t} = ((n^2+1) \cdot i, j \cdot (n^2+1) + t)$$
	
Let's define sets $H$ and $V$ as:
$$H = \{h_{i, j, t} : 1 \le i, j, \le k, 1 \le t \le n^2\}$$
$$V = \{v_{i, j, t} : 1 \le i, j, \le k, 1 \le t \le n^2\}$$
	
Let's define $\epsilon = 0.1$.
For a point $\{x, y\} = p$ we define points
$p^{L} = \{x - \epsilon, y\}$,
$p^{R} = \{x + \epsilon, y\}$,
$p^{U} = \{x, y - \epsilon\}$,
and $p^{D} = \{x, y + \epsilon\}$.

Then we define:
$$\points := H \cup \{p^L : p \in H\} \cup \{p^R : p \in H\}
\cup V \cup \{p^U : p \in V\} \cup \{p^D : p \in V\} $$


\subparagraph{Segments.}
Define horizontal segments.

$$hor_{i, j, t_1, t_2} = (h^R_{i,j,t_1}, h^L_{i, j+1, t_2})$$
$$ver_{i, j, t_1, t_2} = (v^D_{i,j,t_1}, v^U_{i, j+1, t_2})$$

$$horbeg_{i, t} = (h^L_{i, 1, 1}, h^L_{i, 1, t})$$
$$horend_{i, t} = (h^R_{i, n, t}, h^R_{i, n, n^2})$$


$$verbeg_{i, t} = (v^U_{i, 1, 1}, v^U_{i, 1, t})$$
$$verend_{i, t} = (v^D_{i, n, t}, v^D_{i, n, n^2})$$

\begin{eqnarray*}
HOR &= &\{hor_{i, j, t_1, t_2} : 1 \le i \le k, 1 \le j < k,
1 \le t_1, t_2 \le n^2, match_h(t_1, t_2)\} \\
&\cup &\{horbeg_{i,t} : 1 \le i \le k, 1 \le t \le n^2\}
\\
&\cup &\{horend_{i,t} : 1 \le i \le k, 1 \le t \le n^2\}
\end{eqnarray*}

\begin{eqnarray*}
VER &= &\{ver_{i, j, t_1, t_2} : 1 \le i \le k, 1 \le j < k,
1 \le t_1, t_2 \le n^2, match_v(t_1, t_2)\} \\
&\cup &\{verbeg_{i,t} : 1 \le i \le k, 1 \le t \le n^2\}
\\
&\cup &\{verend_{i,t} : 1 \le i \le k, 1 \le t \le n^2\}
\end{eqnarray*}

$$DIAG := \{ (h_{i, j, t}, v_{j, i, t}) :
	1 \le i, j \le k, 1 \le t \le n^2, a_t \in S_{i, j}\}$$

TODO: explain that these segments are in fact diagonal

$$\sets := HOR \cup VER \cup DIAG$$



\begin{lemma}
	If there exists solution for grid tiling,
	then there exists solution for our construction
	using $2(k+1)k + k^2$ segments
	with weight exactly $2k \cdot (k(n^2+1) - 2 - 2\epsilon(k-1))$.
\end{lemma}

\begin{claim}
	If there exists a solution to the grid tiling
	$c_1 \ldots c_k$ and $r_1 \ldots r_k$,
	then there exists a solution covering
	all points
	$$\{h_{i, j, t} : 1 \le i, j \le k, t=(c_i, r_j)\}
	\cup \{v_{i, j, t} : 1 \le i, j \le k, t=(c_j, r_i)\}$$
	
	with segments in $DIAG$
	and the rest in $VER$ or $HOR$
	and has weight $2k \cdot (k(n^2+1) - 2 - 2\epsilon(k-1))$.
\end{claim}

\paragraph{Proof.}
TODO: jakiś prosty z definicji

\begin{lemma}
	If there exists solution for our construction
	using $2(k+1)k + k^2$ segments
	with weight exactly $2k \cdot (k(n^2+1) - 2 - 2\epsilon(k-1))$,
	then there exists a solution for grid tiling
\end{lemma}

\paragraph{Proof.}
This follows from Lemma $\ref{main_soundness_construction}$,
because we just take which points are covered with $DIAG$.

\begin{claim}
\label{guards}
Points $p^{L}, p^R, p^U, p^D$ cannot be covered with $DIAG$.
\end{claim}

\begin{claim}
\label{hor_ver_sound}
Points in $H \cup \{p^L : p \in H\} \cup \{p^R : p \in H\}$
cannot be covered with $VER$.

Points in $V \cup \{p^U : p \in V\} \cup \{p^D : p \in V\} $
cannot be covered with $HOR$.
\end{claim}

\begin{claim}
For given $i, j$ if none of the points $h_{i, j, t}$ ($v_{i, j, t}$)
for $1 \le t \le n^2$ are covered with $DIAG$,
then some spaces between neighbouring points were covered twice.
\end{claim}

\begin{claim}
For given $i, j$ two points $h_{i, j, t_1}, h_{i, j, t_2}$
($v_{i, j, t_1}, v_{i, j, t_2}$)
for $1 \le t_1 < t_2 \le n^2$ are covered with $DIAG$,
then one of them had to be also covered with a segment from $HOR$
($VER$).
\end{claim}
\paragraph{Proof.}
Point $v^L_{i, j, t_2}$ had to be covered with $VER$
from Claims $\ref{guards}$ and $\ref{hor_ver_sound}$.
And every segment in $VER$ covering $v^L_{i, j, t_2}$,
covers also $v^L_{i, j, t_1}$.

\begin{lemma}
	\label{main_soundness_construction}
	If there exists solution for our construction
	with weight at most (exactly)
	$2k \cdot (k(n^2+1) - 2 - 2\epsilon(k-1))$,
	then for every $i, j$
	there must be exactly one $t$ such that $h_{i, j, t}$
	($v_{i, j, t}$) 
	is covered with $DIAG$
	and moreover if $h_{i, j, t_1}$ and $h_{i, j+1, t_2}$
	are uncovered, then $math_h(t_1, t_2)$.
	Analogically for $v$.
\end{lemma}
\paragraph{Proof.}
Only $k^2$ points can be covered only in $DIAG$, the rest
has to be covered with $VER \cup HOR$.
Therefore every result must be at least $ALL\_LINES$ - $2k^2\epsilon$,
because only $2k^2$ spaces of length $\epsilon$
can be uncovered in this axis.

Of course if $h_{i, j, t_1}$ and $h_{i, j+1, t_2}$
are uncovered, then there must exist
a segment in $HOR$ between $h^R_{i, j, t_1}$ and $h^L_{i, j+1, t_2}$,
so $math_h(t_1, t_2)$ must be true.



\subsection{What is missing}
We don't know FPT for axis-pararell segments without $\delta$-extensions.
